Determinism, Complexity, and Predictability in Computer Performance

### 2014-07-14 pecomplexity.tex
address some stuff
edits

\documentclass[%
pre,
reprint,
superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose,
%preprint,
showpacs,
%preprintnumbers,
nofootinbib,
nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-1}
%\documentclass{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parIEEEabrvskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[sort&compress]{natbib}
%\usepackage{caption}
\usepackage{subcaption}
\usepackage{microtype}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{amssymb}
%\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{amsthm}
\usepackage{xspace}
\usepackage{pifont} % check and x marks
\usepackage{color}
\usepackage{listings}
\lstset{language=C}

@@

This paper provides insight into when, why, and how forecast
strategies fail when they are applied to complicated time series. We
conjecture that the inherent complexity of a time series --- which
results from the dimension, nonlinearity, and non-stationarity of the
generating process, as well as from measurement issues like noise,
aggregation, and finite data length --- is both empirically
quantifiable and directly correlated with predictability.  In
particular, we argue that \emph{redundancy} is an effective way to
measure complexity and predictive structure in a time series and that
\emph{weighted permutation entropy} is an effective way to estimate
that complexity.  To validate these conjectures, we study 120
different time-series data sets.  For each time series, we construct
predictions using a wide variety of forecast models, then compare the
accuracy of the predictions with the permutation entropy of that time
series.  We use the results to develop a heuristic that can help
practitioners recognize when a particular prediction method is not
well matched to the task at hand: that is, when the time series has
more predictive structure than that method can capture and exploit.

>>>>

\documentclass[%
pre,
reprint,
superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose,
%preprint,
showpacs,
%preprintnumbers,
nofootinbib,
nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-1}
%\documentclass{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parIEEEabrvskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
%\usepackage{hyperref}
\usepackage[sort&compress]{natbib}
%\usepackage{caption}
\usepackage{subcaption}
\usepackage{microtype}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{amssymb}
%\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{amsthm}
\usepackage{xspace}
\usepackage{pifont} % check and x marks
\usepackage{color}
\usepackage{listings}
\usepackage{rotating}
\usepackage{array,booktabs}
\lstset{language=C}

@@

This paper provides insight into when, why, and how forecast
strategies fail when they are applied to complicated time series. We
conjecture that the inherent complexity of real-world time-series
data---which results from the dimension, nonlinearity, and
non-stationarity of the generating process, as well as from
measurement issues like noise, aggregation, and finite data
length---is both \emph{empirically} quantifiable and directly
correlated with predictability.  In particular, we argue that
\emph{redundancy} is an effective way to measure complexity and
predictive structure in an experimental time series and that
\emph{weighted permutation entropy} is an effective way to estimate
that complexity.  To validate these conjectures, we study 120
different time-series data sets.  For each time series, we construct
predictions using a wide variety of forecast models, then compare the
accuracy of the predictions with the permutation entropy of that time
series.  We use the results to develop a model-free heuristic that can
help practitioners recognize when a particular prediction method is
not well matched to the task at hand: that is, when the time series
has more predictive structure than that method can capture and
exploit.
